Open-source tools in R for landscape ecology

### 2021-01-14 paper/paper.Rmd
Added to the movement ecology paragraph

Because the used modeling approaches are diverse \cite{Hooten2011,Kerr2011,Fletcher2019}, there is also a large number of *R* packages used for SDMs.
Popular approaches and packages include generalized linear models using, e.g., the *stats* \cite{RCoreTeam2019} package; generalized additive models using, e.g., the *mgcv* \cite{Wood2017} or *lme4* \cite{Bates2015} package; classification and regression trees (CART) using, e.g., the *rpart* \cite{Therneau2019}, *randomForest* \cite{Liaw2002} or *ranger* \cite{Wright2017} package or multivariate data analysis using , e.g., the *ade4* \cite{Dray2007} or *vegan* \cite{Oksanen2019} package.
Also, packages specifically designed for SDM exists, including includes the *dismo* \cite{Hijmans2017}, *sdm* \cite{Naimi2016}, *ecospat* \cite{Broennimann2020}, *biomod2* \cite{Thuiller2020}, *PresenceAbsence* \cite{Freeman2008}, or *zoon* \cite{Golding2018} packages.
<!-- These packages [check!] allow the user to create ensemble models of SDMs based on different underlying model structures such as Maxent, regression trees, and generalized linear or additive models.
This is considered the gold standard in contemporary SDM \cite(Araujo2007).-->
Related to SDM, there are several packages to estimate home ranges and habitat selection, including *adehabitatHR* and *adehabitatHS* \cite{Calenge2006} or *amt \cite{Signer2019}*

>>>>

Because the used modeling approaches are diverse \cite{Hooten2011,Kerr2011,Fletcher2019}, there is also a large number of *R* packages used for SDMs.
Popular approaches and packages include generalized linear models using, e.g., the *stats* \cite{RCoreTeam2019} package; generalized additive models using, e.g., the *mgcv* \cite{Wood2017} or *lme4* \cite{Bates2015} package; classification and regression trees (CART) using, e.g., the *rpart* \cite{Therneau2019}, *randomForest* \cite{Liaw2002} or *ranger* \cite{Wright2017} package or multivariate data analysis using , e.g., the *ade4* \cite{Dray2007} or *vegan* \cite{Oksanen2019} package.
Also, packages specifically designed for SDM exists, including includes the *dismo* \cite{Hijmans2017}, *sdm* \cite{Naimi2016}, *ecospat* \cite{Broennimann2020}, *biomod2* \cite{Thuiller2020}, *PresenceAbsence* \cite{Freeman2008}, or *zoon* \cite{Golding2018} packages.
<!-- These packages [check!] allow the user to create ensemble models of SDMs based on different underlying model structures such as Maxent, regression trees, and generalized linear or additive models.
This is considered the gold standard in contemporary SDM \cite(Araujo2007).-->

Related to SDMs, there are growing number of R packages to analyze data from tracked animals, to study their movement characteristics, space use and interaction with other animals and the environment.
These analyses often use results of landscape ecological analyses as explanatory variables to explain variation in space use \cite{signer2015}, behavioral states \cite{langrock2012}, or habitat selection \cite{fieberg2020}.
Widely used R packages include *ctmm* \cite{calabrese2016} and *adehabitatHR* \cite{Calenge2006} for home-range estimation, *moveHMM* \cite{michelot2016} for the classification of behavioral states, and *amt* \cite{Signer2019} for habitat selection.
A recent and very comprehensive overview of R packages for the analysis of animal movement data is given by \cite{joo2020}.

### 2021-01-14 paper/paper.tex
Added to the movement ecology paragraph

Because the used modeling approaches are diverse \cite{Hooten2011,Kerr2011,Fletcher2019}, there is also a large number of \emph{R} packages used for SDMs.
Popular approaches and packages include generalized linear models using, e.g., the \emph{stats} \cite{RCoreTeam2019} package; generalized additive models using, e.g., the \emph{mgcv} \cite{Wood2017} or \emph{lme4} \cite{Bates2015} package; classification and regression trees (CART) using, e.g., the \emph{rpart} \cite{Therneau2019}, \emph{randomForest} \cite{Liaw2002} or \emph{ranger} \cite{Wright2017} package or multivariate data analysis using , e.g., the \emph{ade4} \cite{Dray2007} or \emph{vegan} \cite{Oksanen2019} package.
Also, packages specifically designed for SDM exists, including includes the \emph{dismo} \cite{Hijmans2017}, \emph{sdm} \cite{Naimi2016}, \emph{ecospat} \cite{Broennimann2020}, \emph{biomod2} \cite{Thuiller2020}, \emph{PresenceAbsence} \cite{Freeman2008}, or \emph{zoon} \cite{Golding2018} packages.
Related to SDM, there are several packages to estimate home ranges and habitat selection, including \emph{adehabitatHR} and \emph{adehabitatHS} \cite{Calenge2006} or \emph{amt \cite{Signer2019}}

>>>>

Because the used modeling approaches are diverse \cite{Hooten2011,Kerr2011,Fletcher2019}, there is also a large number of \emph{R} packages used for SDMs.
Popular approaches and packages include generalized linear models using, e.g., the \emph{stats} \cite{RCoreTeam2019} package; generalized additive models using, e.g., the \emph{mgcv} \cite{Wood2017} or \emph{lme4} \cite{Bates2015} package; classification and regression trees (CART) using, e.g., the \emph{rpart} \cite{Therneau2019}, \emph{randomForest} \cite{Liaw2002} or \emph{ranger} \cite{Wright2017} package or multivariate data analysis using , e.g., the \emph{ade4} \cite{Dray2007} or \emph{vegan} \cite{Oksanen2019} package.
Also, packages specifically designed for SDM exists, including includes the \emph{dismo} \cite{Hijmans2017}, \emph{sdm} \cite{Naimi2016}, \emph{ecospat} \cite{Broennimann2020}, \emph{biomod2} \cite{Thuiller2020}, \emph{PresenceAbsence} \cite{Freeman2008}, or \emph{zoon} \cite{Golding2018} packages.

Related to SDMs, there are growing number of R packages to analyze data from tracked animals, to study their movement characteristics, space use and interaction with other animals and the environment.
These analyses often use results of landscape ecological analyses as explanatory variables to explain variation in space use \cite{signer2015}, behavioral states \cite{langrock2012}, or habitat selection \cite{fieberg2020}.
Widely used R packages include \emph{ctmm} \cite{calabrese2016} and \emph{adehabitatHR} \cite{Calenge2006} for home-range estimation, \emph{moveHMM} \cite{michelot2016} for the classification of behavioral states, and \emph{amt} \cite{Signer2019} for habitat selection.
A recent and very comprehensive overview of R packages for the analysis of animal movement data is given by \cite{joo2020}.
